---
title: "COMPSCIX 415.2 Homework 2_herman_andrew"
author: "AndrewHerman"
date: "2/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(tidyverse)
```
# Section 3.2.4: all exercises
1) Run ggplot(data = mpg). What do you see?  
*I see a blank canvas with nothing on it.*

```{r}
ggplot(data = mpg)
```

2) How many rows are in mpg? How many columns?  

*There are this many rows in mpg:*  
```{r}
nrow(mpg)
```
*There are this many columns:*  
```{r}
ncol(mpg)
```

3) What does the drv variable describe? Read the help for ?mpg to find out.  
*The drv variable describes whether the care is front wheel drive, rear wheel drive, or four wheel drive.*

```{r}
?mpg
```


4) Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data = mpg, aes(x = hwy, y = cyl)) + 
  geom_point()
```

5) What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r}
ggplot(data = mpg, aes(x = class, y = drv)) + 
  geom_point()
```

*This plot is not useful because it only shows which drive types are applicable to each car class.  A chart is not necessary to view this data since it is very basic and essentially binary.  A table would be more concise and easier to understand.*

# Section 3.3.1: all exercises  

1) What’s gone wrong with the code? Why are the points not blue?  
*The points are not blue because the color argument needs to be set outside of the 'aes' function.  The correct code is below:*  
```{r}
ggplot(data = mpg) + 
geom_point(mapping = aes(x = displ, y = hwy), color = 'blue')
```

2) Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?  
*The below code will display the first 6 lines of data, as well as the headers.  If there is a '<chr>' at the top of the column, this means that the column is categorical.  All other columns are continuous.*  
*Categorical: manufacturer, model, trans, drv, fl, and class*  
*Continuous: displ, year, cyl, cty, hwy*

```{r}
head(mpg)
```

3) Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?  

*When year is mapped to color, the coninuous variables are displayed on a color scale.*  
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, color = year)) +
  geom_point()
```

*When year is mapped to size, the coninuous variables are displayed depending on their size.*  
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, size = year)) +
  geom_point()
```

*When year is mapped to shape, an error was returned, because shape does not have an associated scale.  Note: the code below is commented out due to the erroe.*  
```{r}
# ggplot(data = mpg, aes(x = displ, y = hwy, shape = year)) +
  # geom_point()
```

4) What happens if you map the same variable to multiple aesthetics?  
*If the same variable is applied to multiple aesthetics, a chart is still produced, however this is not a useful chart.  In the below example, hwy is applied to the y axis, size, and color.  The chart now shows the same information three times, which makes the visualization more difficult to interpret.*

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, size = hwy, color = hwy)) +
  geom_point()
```

5) What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)  

*The stroke aesthetic changes the thickness of the shape's border.  The different shapes can be found [here](http://www.cookbook-r.com/Graphs/Shapes_and_line_types/figure/unnamed-chunk-2-1.png).*
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, size = year)) +
  geom_point(stroke = 3, shape = 5)
```
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, size = year)) +
  geom_point(stroke = 8, shape = 5)
```


6) What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.  

*The size aesthetic now representants an IF statement, where the false values are smaller.  In this case, if the engine displacement is less than 5, there will be a big circle.*

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, size = displ < 5)) +
  geom_point()
```

# Section 3.5.1: #4 and #5 only

4) Take the first faceted plot in this section:

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

*The advantages are that each variable is on a different chart allowing the reader to both easily spot difference between the charts, but analyze the single variable on its own.  The disadvantage to this is that it does make a side by side comparison a bit more difficult and less straight forward.  Additoinally, outliers may be more easily overlooked since they would not necessarily standout.  A larger dataset would make faceting more beneficial, because a single chart with a large number of datapoints would be very difficult to read and interpret.*

5) Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

*nrow counts the number of rows in a dataset, while ncol counts the numbers of columns.  Some other options to control the layout of panels include (but are not limited to): shrink, labeller, switch, and drop.*

*facet_grid() does not have nrow nor ncol arguments because the panels are already defined by row and column faceting variables.*

# Section 3.6.1: #1-5. Extra Credit: Do #6  

1) What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

*For a line chart, I would use the smoooth geom.  For a boxpliot, I would use the boxplot geom. For a histogram, I would use the histogram geom. For an area chart, I would use the area geom.*

2) Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

*I predict that the smooth line will be based on the hwy mileage, while the points will represent the drv.*

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE, show.legend = FALSE)
```

3) What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

*In the above chart, the show.legend = false removed the line from the legend, but the legend still remained for the colors and dots.  This is because I placed the show.legend code as part of the geom_smooth code.  The legend was suppressed earlier in the chapter so that the shape/size of the graph would not be impacted by the legend*

4) What does the se argument to geom_smooth() do?

*Per the help screen (i.e. ?geom_smooth), the geom_smooth "Aids the eye in seeing patterns in the presence of overplotting"*

5) Will these two graphs look different? Why/why not?

*These two graphs should not look different because the code is doing the same thing.  On the top graph, the constant arguments are declared for the whole ggplot, while the bottom graph simply repeats the same code, but for both geoms.  The first chart is cleaner code, but the second chart could allow for greater customization if desired.*

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

6) Recreate the R code necessary to generate the following graphs.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() + 
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, line = drv)) +
  geom_point() + 
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() + 
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv)) + 
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy), show.legend = TRUE) +
  geom_point(aes(color = drv)) + 
  geom_smooth(se = FALSE, aes(line = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy), show.legend = TRUE) +
  geom_point(aes(color = drv), size = 5, outline = "white")
```

# Section 3.7.1: #2 only

2) What does geom_col() do? How is it different to geom_bar()?

*According to the ?geom_col help screen, "geom_bar() makes the height of the bar proportional to the number of cases in each group (or if the weight aesthetic is supplied, the sum of the weights). If you want the heights of the bars to represent values in the data, use geom_col()"*

# Answer these questions:

1) Look at the data graphics at the following link: [What is a Data Scientist](https://static.guim.co.uk/sys-images/Guardian/Pix/pictures/2012/3/2/1330695817953/EMC2-graphic-on-data-scie-008.jpg). Please briefly critique the designer’s choices. What works? What doesn’t work? What would you have done differently?

*A) I would change the proportion of the image to fit cleanly on a 16:9 display*  
*B) I would not have the graph paper style in the background.  It adds noise to an already busy visual.*  
*C) There are lots of percentages, but I know nothing about the survey.  How many people were asked?  Who are they?  What are the biases? etc.*  
*D) This is a poorly written statement "Over 2/3 beleive demand for talent will outpace the supply of data scientists".  Over 2/3 what?  Monkeys?  Bankers?  Data scientists?*  
*E) This is another bad statement "Only 12% see today's BI professional as the best srouce for new data scientists".  What does BI stand for?  Business Intelligence?  Busy Intern?  Best Intellect?*  
*F) "The best source of new data science talent is:".  There is a ton of screen space.  The author could add a decimal place to these numbers and make them add up to 100%.*  
*G) Halfway through the doc, there are way too many font sizes, colors, transparancies all caps, no caps, proper caps.  The text is fatiguing to read.*  
*F) Overall there is not a cohesive storyline to this document.  I'm not sure what argument the is being made*  

