---
title: "andrew_herman_final"
author: "AndrewHerman"
date: "3/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(nycflights13)
library(dplyr)
library(tibble)
library(mosaicData)
library(broom)
library(rvest)
library(jsonlite)
library(rgdal)
library(ggmap)
library(leaflet)
library(shiny)
```

# Exercise 1 - Sampling Distributions, Functions and For Loops (10 points)  

## Step 1  
Write an R function that does the following:  
-Takes a sample of size samp_size from this exponential distribution (samp_size is an input parameter for the function)  
-Calculates the mean of that sample  
-Calculates the standard deviation of that sample  
-Returns the calculated mean and standard deviation as a list  

```{r}

samp_fun <- function(samp_size, samp_rate) {
  
  samp1 <- rexp(n = samp_size, rate = samp_rate)
  
  samp_avg <-mean(samp1)
  samp_std_dev <- sd(samp1)
  
  stats <- list(samp_avg = samp_avg, samp_std_dev = samp_std_dev)
  return(stats)
}

samp_fun(100, 1/10000)
```

## Step 2
Then write a loop that does this:  
Runs the above function 1000 times, with samp_size = 50 and samp_rate = 1/10000
Saves all of the sample means in a vector called sample_means, and all of the sample standard deviations in a vector called sample_sds

```{r}
samp_fun <- function(samp_size, samp_rate) {

  x <- 1000
  sample_means <- rep(NA, x)
  sample_sds <- rep(NA, x)

  for (i in 1:x) {
    samp1 <- rexp(n = samp_size, rate = samp_rate)
    
    samp_avg <- mean(samp1)
    samp_std_dev <- sd(samp1)
  
    sample_means[i] <- samp_avg
    sample_sds[i] <- samp_std_dev

  }

  results <- list(sample_means = sample_means, sample_sds = sample_sds)
  return(results)  

}
samp_fun(50, 1/10000)
```


## Step 3  

### Plot your sample means as a histogram  

```{r}
step3data <- samp_fun(50, 1/10000)

sample_means_L <- step3data[1]
sample_sds_L <- step3data[2]

sample_means_df <- data.frame("sample_means" = step3data[1])
sample_sds_df <- data.frame("sample_sds" = step3data[2])

ggplot(data = sample_means_df, aes(x = sample_means)) +
  geom_histogram()

```


### Output the standard deviation of your sample means  

```{r}
samp_mean_stdev <- sd(sample_means_df$sample_means)
samp_mean_stdev
```

### Calculate the theoretical standard error (σ=10000, n = sample size)  



### Calculate the mean of the sample standard deviations and use this to calculate the empirical standard error  





## Step 4  
Repeat STEP 2 and STEP 3 using a sample size of 5000.



```{r}

```



# Exercise 2 - Linear Regression (5 points)  

## Load the train.csv dataset into R and fit a regression model with:
y = SalePrice  
Features: LotArea, OverallQual, and ExterQual  

```{r}
train <- read_csv("/Users/aherman/Downloads/train.csv")

train_reg <- lm(SalePrice ~ LotArea + OverallQual + ExterQual, data = train)


```

## Use the broom package to output the coefficients and the R-squared

```{r}
glance(train_reg)
tidy(train_reg)
```

## Interpret the coefficient on LotArea  
LotArea is statistically significant.  For every squre foot increase in Lot size, the average sale prie increaed by $1.45, assuming all else is equal.  

## Interpret the coefficient on ExterQualGd
The mean difference in sale price for a Good exterior quality and an Excellent exterior quality is -71,529 dollars.  In other words, with all else equal, going from an Excellent to a Good will result in a sales price decrease of 71,529 dollars.

## Compare this model to the model we fit in HW 7 with GrLivArea, OverallQual, Neighborhood. Which is the better fitting model?  
The R-squared from the model in HW 7 is equal to .78.  The R-squared in this model is .69, which indicates this model is less accurate.

# Exercise 3 - AB Testing (5 points)  
We want to perform an AB test on this data to see if the conversion rates are different for the two versions of the webpage.

Answer these questions:

## What proportion of visitors converted for each version of the webpage?

```{r}
testdata <- read_csv("/Users/aherman/Downloads/ab_test_data.csv")

proportions <- testdata %>%
  group_by(version) %>%
  summarise(tot_conversions = sum(conversion), prop_conversion = mean(conversion))
proportions 

```

4.15% of visitors converted for version A, while 10% of visitors converted for version B.

## Perform the AB test in R. What is the p-value for the AB test (hypothesis test of proportions)?

```{r}
true_a <- .0415
true_b <- .1
n_a <- 1000
n_b <- 1000

set.seed(10)
samp_a <- rbinom(n = 1, size = n_a, prob = true_a)
samp_b <- rbinom(n = 1, size = n_b, prob = true_b)

two_prop_test <- prop.test(c(samp_a, samp_b), c(1000, 1000))
two_prop_test$p.value
```
The p-value is less than .05, therefore we know that the conversion rates between version A and B are significantly different.

